---
title: "CCtunnelAnalysis"
author: "Sean Liu"
date: "2023-11-20"
output:
  rmdformats::readthedown:
    self_contained: true
    code_download: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    df_print: paged
    hightlight: tango
    code_folding: hide
    fig_width: 10
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# load packages
library(data.table)
library(ggplot2)
library(ggpubr)
library(readxl)
library(reshape2)
library(dplyr)
library(rstatix)
library(Rmisc)
library(lattice)
library(plyr)
library(ggsci)
library(emmeans)
library(envalysis)
library(gridExtra)
library(ppcor)
library(psych)
library(Hmisc)
library(kableExtra)
library(corrplot)
library(ppcor)
library(patchwork)
library(gridExtra)
library(grid)
library(ggsignif)
library(writexl)
library(stringr)
library(RColorBrewer)
library(splitstackshape)
library(lme4)
library(sjstats)
library(lmerTest)
library(MuMIn)
library(reshape)
library(tidyr)
library(BayesFactor)


options(digits = 4) 
options(warn = -1)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


# Data exclusion

Load pre-processed data from Pavlovia and do data cleaning in accordance
with criterion:\
<!-- (1)Subjects whose error rates exceed mean ± 3 SD of the group -->
<!-- (2)Subjects whose RT variation exceed mean ± 3 SD of the group -->
(1)Individual trials with RT shorter than 200ms or exceeding the outlier
threshold of remaining trials(3SD)\
(2)Subjects whose with average accuracy lower than 90% are cut off\
(3)RTs are only analyzed for correct responses\
(4)Bin 24 blocks in main experiment into 6 epochs to increase
statistical power\

Experimental design: 16 trials in each block, including 8 repeated
config + 8 novel config; 25 blocks in total, including 20 training
blocks + 4 testing blocks + 1 recognition task.

```{r clean data}
rm(list = ls())
# read excel sheet with name
rawdataall <- read_excel("AllData_exp2.xlsx", sheet="Exp2RawData")
setDT(rawdataall)

# remove practice phase
rawdataall <- rawdataall[phase %in% c("training","testing","rt")]

# set names for presentation
setnames(rawdataall,"participant","Sub")
setnames(rawdataall,"CCCondition","Configuration")
setnames(rawdataall,"key_respMain.corr","Accuracy")
setnames(rawdataall,"key_respMain.rt","RT")
setnames(rawdataall,"phase","Phase")
setnames(rawdataall,"block","Block")
setnames(rawdataall,"targcell","Cell")

# convert into numeric
rawdataall$RT = as.numeric(rawdataall$RT)
rawdataall$Accuracy = as.numeric(rawdataall$Accuracy)
initial_count <- nrow(rawdataall)
###########################################################
# filter3: only accurate responses(get rid of table format)
tempdata = rawdataall
tempdata[,filter:=1][Accuracy==0,filter:=0]
excluded_count_acc <- sum(tempdata$filter == 0)
cat("Number of excluded trials based on Acc:", excluded_count_acc, "\n")
###########################################################


## filter1: Individual trials remove based on RT
lowbound <- 0.2
# function to recursively remove outliers(typically 1 round only)
remove_RToutliers <- function(data, column_name, threshold_multiplier = 3, max_iterations = 1) {
  for (iteration in 1:max_iterations) {
    # Calculate mean and standard deviation
    mean_val <- mean(data[[column_name]])
    sd_val <- sd(data[[column_name]])
    
    # Set the threshold for outliers
    lower_threshold <- mean_val - threshold_multiplier * sd_val
    upper_threshold <- mean_val + threshold_multiplier * sd_val
    
    # Remove outliers
    data <- data %>% filter((data[[column_name]] > lower_threshold) & (data[[column_name]] < upper_threshold))
  }
  return(data)
}

# Remove outliers in RT for main exp but not recognition task -- continue on filter1
setDT(rawdataall)
tempdata <- rawdataall %>% filter((RT >= lowbound & Phase %in% c("training", "testing")) | !(Phase %in% c("training", "testing")))

tempdata <- tempdata %>% filter(Phase != "rt")
tempdata <- tempdata %>%
  group_by(Sub) %>%
  group_modify(~ remove_RToutliers(.x, "RT"))
tempdata <- bind_rows(tempdata, rawdataall %>% filter(Phase == "rt"))

excluded_count_rt <- initial_count - nrow(tempdata)
cat("Number of excluded trials based on RT:", excluded_count_rt, "\n")


# ## filter2: Individual blocks remove based on Acc(90% cutoff)
# accCutoff_data <- tempdata %>%
#   filter(Phase %in% c("training", "testing"))
# 
# # Calculate mean accuracy
# sub_means <- accCutoff_data %>%
#   group_by(Sub, Configuration, Phase) %>%
#   summarise(mean_accuracy = mean(Accuracy)) %>%
#   group_by(Sub) %>%
#   summarise(mean_accuracy = mean(mean_accuracy))
# 
# 
# # Identify 'Sub' with mean accuracy less than 0.9
# subs_to_block <- sub_means %>%
#   filter(mean_accuracy < 0.9) %>%
#   pull(Sub)
# 
# # Count and print the number of 'Sub' being ruled out
# num_subs_ruled_out <- length(subs_to_block)
# cat(paste("Number of Sub ruled out due to mean accuracy < 0.9 under both training and testing phases:", num_subs_ruled_out, "\n"))
# 
# # Filter out the data for the 'Sub' that needs to be blocked
# cleaned_data <- data %>%
#   filter(!(Sub %in% subs_to_block))


##filter4: See in RT ANOVA
```

# Accuracy

Check and basically plot: 1.Overall statistics for accuracy 2.two-way
within-subject ANOVA on acc(If any speed-acc trade-offs with regard to
configuration)

## Averaged accuracy plotting

```{r bar plot}
# Plot 1
# ACC by config
tempdata <- as.data.table(tempdata)
accuracylong <- tempdata[, .(Accuracy=mean(Accuracy)),keyby=.(Sub, Configuration, Block, Phase)]
accuracylong$Sub = as.factor(accuracylong$Sub)
accuracylong$Configuration = as.factor(accuracylong$Configuration)
accuracylong$Block <- ifelse(accuracylong$Phase == "testing", accuracylong$Block + 20, accuracylong$Block)
accuracylong <- accuracylong %>% filter(Phase != "rt")

###################################################
## filter2: 90% cutoff for each Sub
mean_accuracy <- accuracylong %>%
  group_by(Sub) %>%
  mutate(mean_acc = mean(Accuracy))

Subs_cutoff <- mean_accuracy %>%
  distinct(Sub, mean_acc) %>%
  filter(mean_acc < 0.9) %>%
  pull(Sub)

num_subs_ruled_out <- length(Subs_cutoff)
cat(paste("Number of excluded subjects based on Accuracy:", num_subs_ruled_out, "\n"))
###################################################

# Mean and within-subject SE
dfwc <- summarySEwithin(accuracylong, measurevar="Accuracy", withinvars=c("Configuration"),
                        idvar="Sub", na.rm=FALSE, conf.interval=.95)
# bar plot for acc
b_acc <- dfwc %>%
  ggplot(aes(x=factor(Configuration,levels=c("novel","repeated")), 
             y=Accuracy*100, 
             fill=factor(Configuration,levels=c("novel","repeated")))) +
  geom_bar(position=position_dodge(.9), fill = c("novel" = "white", "repeated" = "grey"), width=.8, colour="black", stat="identity") +
  geom_errorbar(position=position_dodge(.9), width=.2, aes(ymin=100*(Accuracy-se), ymax=100*(Accuracy+se))) +
  coord_cartesian(ylim=c(90,100)) +
  xlab("Target's Configuration") +
  ylab("Accuracy (%)") +
  theme_publish() +
  theme(legend.position="top",
        aspect.ratio = 1.5) +
  scale_fill_grey(start = 0.2, end = 0.6) + 
  guides(fill=guide_legend(title="Configuration",nrow=1,byrow=TRUE))
print(b_acc)


```

## Accuracy ANOVA

```{r acc ANOVA}
##filter4: Binned into epochs
accuracylong <- as.data.table(accuracylong)
accuracylong[, Block := as.numeric(Block)]
accuracylong[, Epoch := cut(Block, breaks = c(0, 4, 8, 12, 16, 20, 25),
                        labels = c(1, 2, 3, 4, 5, 6), include.lowest = FALSE)]
accuracylong[, Epoch := as.numeric(Epoch)]
accuracylong$Epoch <- as.factor(accuracylong$Epoch)
accuracylong <- accuracylong[Epoch %in% c(1, 2, 3, 4, 5)]
accuracylong <- accuracylong[, .(Accuracy = mean(Accuracy)),
                     keyby = .(Sub,Configuration)]

# repeated ANOVA: Block * Configuration
options(contrasts=c('contr.sum','contr.poly'))
res.aov <- anova_test(
  data = accuracylong,
  dv = Accuracy,
  wid = Sub,
  within = Configuration,
  type = 3,
  effect.size = "pes"
)
# print ANOVA table
get_anova_table(res.aov, 
                correction="none") %>%
  kbl(caption = "Accuracy: ANOVA on target's configuration * epoch") %>%
  row_spec(0,bold=TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial",font_size = 16)

accuracy_wide <- accuracylong %>%
  pivot_wider(names_from = Configuration, values_from = Accuracy)
t_test_result <- t.test(accuracy_wide$novel, accuracy_wide$repeated, paired = TRUE)
formatted_t_test_result <- lapply(t_test_result, function(x) if(is.numeric(x)) sprintf("%.3f", x) else x)

# Print the formatted result
print(abs(t_test_result$statistic))
cohen_d <- abs(t_test_result$statistic)/sqrt(length(unique(accuracylong$Sub[!is.na(accuracylong$Sub)])))
cat("Cohen's d:", cohen_d, "\n")

# Mean and SE
result <- accuracylong[, .(MeanAccuracy = mean(Accuracy),
                           SE = sd(Accuracy)), 
                       by = Configuration]

print(result)
```

Overall accuracy is high for both configuration across epochs. The repeated ANOVA on accuracies indicates no potential speed-accuracy trade-offs regarding configuration type. However, the main effect of epoch is significant, showing general improvement over time.

# Search Time

Check and basically plot: 1.Overall statistics for RT by binned epoch
2.two-way within-subject ANOVA on RT

## Averaged RT plots

```{r RT plot}
##filter4: Correct trials only
# RT by Config across blocks(correct only)
RTlong <- tempdata[filter==1,
                     .(RT = mean(RT)),
                     keyby = .(Sub,Phase,Configuration,Block, Accuracy)]
RTlong$Sub <- as.factor(RTlong$Sub)
RTlong$Phase <- as.factor(RTlong$Phase)
RTlong$Configuration <- as.factor(RTlong$Configuration)
RTlong$Block <- as.factor(RTlong$Block)
RTlong <- RTlong %>% filter(Phase != "rt")

## Histogram of Epoch 6
RT_specific_blocks <- RTlong[Phase == "testing"]
dfwc_testing <- summarySEwithin(RT_specific_blocks, measurevar = "RT", withinvars = c("Configuration", "Block"), idvar = "Sub", na.rm = FALSE, conf.interval = 0.95)
setDT(dfwc_testing)

# Line plot with connected dots for the within-subject SE of each configuration for each Block
(line_plot_se_testing <- ggplot(dfwc_testing, aes(x = as.factor(Block), y = RT * 1000, color = Configuration)) +
  geom_line(aes(group = Configuration), size = 1) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = (RT - se) * 1000, ymax = (RT + se) * 1000), width = 0.3) +
  scale_color_grey(start = 0.2, end = 0.6) +  # Set color range
  scale_y_continuous(breaks = seq(1300, 3400, by=300), limits=c(1300, 3400)) +
  labs(x = "Testing Block", y = "RT (ms)", title = "Exp 2") +
  geom_vline(xintercept = 5.4, linetype = "dotted") +
  labs(color = "Configuration", shape = "Configuration") +
  theme_publish() +
  theme(
    legend.position = "bottom",
    aspect.ratio = 0.8,
    legend.text = element_text(size = 14),  # legend font size
    legend.title = element_text(size = 14),  # legend title
    axis.text = element_text(size = 14),  # axis text size
    axis.title = element_text(size = 16),  # axis title size
    plot.title = element_text(size = 18, hjust = 0.5)  # plot title size
  ))

# Save the line graph plot with connected dots and shorter aspect ratio for the specific Blocks
ggsave("Exp2RT_Testing.jpg", line_plot_se_testing + theme(plot.background = element_rect(fill = "white")))


RT_last3blocks <- RT_specific_blocks%>%
  filter(Block %in% c("2", "3", "4"))
part.aov <- anova_test(
  data = RT_last3blocks,
  dv = RT,
  wid = Sub,
  within = c(Configuration,Block),
  type = 3,
  effect.size = "pes"
)

##########################################


##filter4: Binned into epochs
RTlong <- as.data.table(RTlong)
RTlong[, Block := as.numeric(Block)]
RTlong$Block <- ifelse(RTlong$Phase == "testing", RTlong$Block + 20, RTlong$Block)
RTlong$Block <- RTlong$Block - 1
RTlong[, Epoch := cut(Block, breaks = c(0, 4, 8, 12, 16, 20, 25),
                        labels = c(1, 2, 3, 4, 5, 6), include.lowest = FALSE)]
RTlong[, Epoch := as.numeric(Epoch)]
RTlong$Epoch <- as.factor(RTlong$Epoch)

# within-subject SE
dfwc <- summarySEwithin(RTlong, measurevar="RT", withinvars=c("Phase","Configuration","Epoch"),
                        idvar="Sub", na.rm=FALSE, conf.interval=.95)
setDT(dfwc)

RT_all_exp <- dfwc %>% 
  ggplot(aes(x=Epoch, y=RT*1000, 
             colour=factor(Configuration,levels=c("novel","repeated")),
             group=factor(Configuration,levels=c("novel","repeated")))) +
  geom_line(size=1) +
  geom_point(aes(shape=factor(Configuration,levels=c("novel","repeated")),
                 color=factor(Configuration,levels=c("novel","repeated"))),size=2.5) + 
  scale_y_continuous(breaks = seq(1300, 3400, by=300), limits=c(1300, 3400)) +
  geom_errorbar(width=.3, aes(ymin=(RT-se)*1000, ymax=(RT+se)*1000)) +
  xlab("Epoch") +
  ylab("RT (ms)") +
  scale_color_grey(start = 0.2, end = 0.6) +
  theme_publish() +
  geom_vline(xintercept = 5.4, linetype="dotted") +
  annotate("text", x=3, y=3300, label= "training", size=15/.pt) + 
  annotate("text", x=6, y=3300, label= "testing", size=15/.pt) +
  labs(color = "Configuration", shape = "Configuration")+
  labs(title = "Exp 2") +
  theme(legend.position="top",
        aspect.ratio = 0.8)+
  theme(legend.text = element_text(size = 14), # legend font size
        legend.title=element_text(size=14), # legend title
        axis.text = element_text(size = 14), # axis text size
        axis.title = element_text(size = 16), # axis title size
        plot.title = element_text(size = 18,hjust = 0.5), # plot title size
  ) 



plot <- ggarrange(RT_all_exp, nrow=1,common.legend = TRUE,legend="bottom")
annotate_figure(plot, top = text_grob("",
                                      color = "black", face = "bold", size = 16))
ggsave("Exp2RT.png", plot + theme_minimal() + theme(plot.background = element_rect(fill = "white")))
```


## RT ANOVA

### Configuration \* Epoch

Experimental blocks are aggregated to 6 epochs, each containing 4
blocks.

```{r RT ANOVA by epoch}
# RT by Config across blocks(correct only)
tempdata <- as.data.table(tempdata)
write.csv(tempdata, file = "RT_exp2.csv")
RTexp <- tempdata[filter==1,
                     .(RT = mean(RT)),
                     keyby = .(Sub,Phase,Configuration,Block)]
RTexp$Sub <- as.factor(RTexp$Sub)
RTexp$Phase <- as.factor(RTexp$Phase)
RTexp$Configuration <- as.factor(RTexp$Configuration)
RTexp$Block <- ifelse(RTexp$Phase == "testing", RTexp$Block + 20, RTexp$Block)
RTexp <- RTexp %>% filter(Phase != "rt")
RTexp$Block <- as.factor(RTexp$Block)

##filter4: Binned into epochs
RTexp <- as.data.table(RTexp)
RTexp[, Block := as.numeric(Block)]
RTexp[, Epoch := cut(Block, breaks = c(0, 4, 8, 12, 16, 20, 25),
                        labels = c(1, 2, 3, 4, 5, 6), include.lowest = FALSE)]
RTexp[, Epoch := as.numeric(Epoch)]
RTexp$Epoch <- as.factor(RTexp$Epoch)
RTexp <- RTexp[, .(RT = mean(RT)),
                     keyby = .(Sub,Configuration,Epoch)]
write.csv(RTexp, file = "Exp2AData.csv")


# repeated ANOVA for RT for main experiment
options(contrasts=c('contr.sum','contr.poly'))
res.aov <- anova_test(
  data = RTexp,
  dv = RT,
  wid = Sub,
  within = c(Configuration,Epoch),
  type = 3,
  effect.size = "pes"
)

# simple effect test
model <- aov(RT ~ Configuration * Epoch + Error(Sub), data = RTexp)
emm <- emmeans(model, ~ Configuration | Epoch)
comparisons <- pairs(emm)
print(comparisons)

# print ANOVA table for main experiment
get_anova_table(res.aov, 
                correction="none") %>%
  kbl(caption = "RT: ANOVA on target's configuration * epoch") %>%
  row_spec(0,bold=TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial",font_size = 16)


```



### ANOVA for Training Phase Only
```{r RT ANOVA by subset epoch}
# Filter data for epochs 1 to 5
RTexp_subset <- RTexp[Epoch %in% c(1, 2, 3, 4, 5)]

# Repeated ANOVA for RT for epochs 1 to 5
res.aov_subset <- anova_test(
  data = RTexp_subset,
  dv = RT,
  wid = Sub,
  within = c(Configuration, Epoch),
  type = 3,
  effect.size = "pes"
)


# Bayes Factor Analysis
bf <- anovaBF(RT ~ Configuration, data = RTexp_subset)
bf


# Simple effect test for epochs 1 to 5
model_subset <- aov(RT ~ Configuration * Epoch + Error(Sub), data = RTexp_subset)

emm_subset1 <- emmeans(model_subset, ~ Configuration | Epoch)
comparisons_subset1 <- pairs(emm_subset1)
print(comparisons_subset1)

emm_subset2 <- emmeans(model_subset, ~ Epoch | Configuration)
comparisons_subset2 <- pairs(emm_subset2)
print(comparisons_subset2)

# Print ANOVA table for epochs 1 to 5
get_anova_table(res.aov_subset, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration * epoch (Epochs 1-5)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)
```



### ANOVA for 5th Epoch
```{r ANOVA final training phase}
# Filter data for the 6th epoch
RTexp_epoch5 <- RTexp[Epoch == 5]

# ANOVA for RT for the 6th epoch (main effect of Configuration)
model_epoch5 <- anova_test(
  data = RTexp_epoch5,
  dv = RT,
  wid = Sub,
  within = Configuration,
  type = 3,
  effect.size = "pes"
)


# Cohen's d
repeated_5 <- RTexp_epoch5$RT[RTexp_epoch5$Configuration == "repeated"]
novel_5 <- RTexp_epoch5$RT[RTexp_epoch5$Configuration == "novel"]
cat("RT saving (s):", mean(novel_5) - mean(repeated_5), "\n")
cat("saving percentage:", (mean(novel_5) - mean(repeated_5))/mean(novel_5), "\n")


# Print ANOVA table for the 5th epoch with PES
get_anova_table(model_epoch5, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration (Epoch 5)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)

# t test
RTepoch5_wide <- RTexp_epoch5 %>%
  pivot_wider(names_from = Configuration, values_from = RT)
t_test_result <- t.test(RTepoch5_wide$repeated, RTepoch5_wide$novel, paired = TRUE)
formatted_t_test_result <- lapply(t_test_result, function(x) if(is.numeric(x)) sprintf("%.3f", x) else x)

# Print the formatted result
print(abs(t_test_result$statistic))

cohen_d <- t_test_result$statistic/sqrt(length(unique(RTexp_epoch5$Sub[!is.na(RTexp_epoch5$Sub)])))
cat("Cohen's d:", cohen_d, "\n")
```



## ANOVA for Testing Phase
```{r ANOVA raw test phase}
# Filter data for the 6th epoch
RTexp_epoch6 <- RTexp[Epoch == 6]

# ANOVA for RT for the 6th epoch (main effect of Configuration)
model_epoch6 <- anova_test(
  data = RTexp_epoch6,
  dv = RT,
  wid = Sub,
  within = Configuration,
  type = 3,
  effect.size = "pes"
)


# Cohen's d
repeated_6 <- RTexp_epoch6$RT[RTexp_epoch6$Configuration == "repeated"]
novel_6 <- RTexp_epoch6$RT[RTexp_epoch6$Configuration == "novel"]
cat("RT saving:", mean(novel_6) - mean(repeated_6), "\n")
cat("saving percentage:", (mean(novel_6) - mean(repeated_6))/mean(novel_6), "\n")


# Print ANOVA table for the 6th epoch with PES
get_anova_table(model_epoch6, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration (Epoch 6)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)

# t test
RTepoch6_wide <- RTexp_epoch6 %>%
  pivot_wider(names_from = Configuration, values_from = RT)
t_test_result <- t.test(RTepoch6_wide$repeated, RTepoch6_wide$novel, paired = TRUE)
formatted_t_test_result <- lapply(t_test_result, function(x) if(is.numeric(x)) sprintf("%.3f", x) else x)

# Print the formatted result
print(t_test_result$statistic)

cohen_d <- t_test_result$statistic/sqrt(length(unique(RTexp_epoch6$Sub[!is.na(RTexp_epoch6$Sub)])))
cat("Cohen's d:", cohen_d, "\n")

```
## ANOVA For Testing Phase by Block
```{r 6th epoch by block}
# RT by Config across blocks(correct only)
tempdata <- as.data.table(tempdata)
write.csv(tempdata, file = "RT_exp2.csv")
RTblock <- tempdata[filter==1,
                     .(RT = mean(RT)),
                     keyby = .(Sub,Phase,Configuration,Block)]
RTblock$Sub <- as.factor(RTblock$Sub)
RTblock$Phase <- as.factor(RTblock$Phase)
RTblock$Configuration <- as.factor(RTblock$Configuration)
RTblock$Block <- ifelse(RTblock$Phase == "testing", RTblock$Block + 20, RTblock$Block)
RTblock <- RTblock %>% filter(Phase != "rt")
RTblock$Block <- as.factor(RTblock$Block)


##filter4: Binned into epochs
RTblock <- as.data.table(RTblock)
RTblock[, Block := as.numeric(Block)]
RTblock <- RTblock[, .(RT = mean(RT)),
                     keyby = .(Sub,Configuration,Block)]

RTblock_epoch6 <- RTblock[Block %in% c(21, 22, 23, 24)] 
RTblock_epoch6$Block <- as.factor(RTblock_epoch6$Block)


# repeated ANOVA for RT for main experiment
# repeated ANOVA for RT for main experiment
options(contrasts=c('contr.sum','contr.poly'))
res.aov <- anova_test(
  data = RTblock_epoch6,
  dv = RT,
  wid = Sub,
  within = c(Configuration,Block),
  type = 3,
  effect.size = "pes"
)

# simple effect test
model <- aov(RT ~ Configuration * Block + Error(Sub), data = RTblock_epoch6)
emm <- emmeans(model, ~ Configuration | Block)
comparisons <- pairs(emm)
print(comparisons)

# print ANOVA table for main experiment
get_anova_table(res.aov, 
                correction="none") %>%
  kbl(caption = "RT: ANOVA on target's configuration * block in training phase") %>%
  row_spec(0,bold=TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial",font_size = 16)

```


# Awareness

To study the effect of explicit awareness of scene configuration
repetition on RT, we convert the forced-choice recognition responses
into d-prime.

## Accuracy for each subject

```{r awareness check}
# RT accuracy for each subject
rtdata <- tempdata %>% filter(Phase == "rt")
rtdata <- as.data.table(rtdata)
accuracyAll <- rtdata[, .(Accuracy=mean(Accuracy)),keyby=.(Sub)]

# print table
accuracyAll %>%
  kbl(caption = "Accuracy of Recognition Task in each subject") %>%
  row_spec(0,bold=TRUE) %>%   kable_classic(full_width = F, html_font = "Arial",font_size = 16)


# t-test
mean(accuracyAll$Accuracy)
chance_level <- 0.5
t.test(accuracyAll$Accuracy, mu = chance_level)

```

## d-prime for each subject

```{r d-prime calculation}
rtdata <- rtdata %>%
  mutate(
    hit = as.numeric(Accuracy == 1 & key_respMain.keys == "y"),
    false_alarm = as.numeric(Accuracy == 0 & key_respMain.keys == "y")
  )
hitAll <- rtdata[, .(hit=mean(hit)),keyby=.(Sub)]
faAll <- rtdata[, .(false_alarm=mean(false_alarm)),keyby=.(Sub)]
dAll <- merge(hitAll, faAll, by = "Sub", all = TRUE)

dAll <- as.data.table(dAll)
dAll[, dprime := qnorm(hit) - qnorm(false_alarm)]
dAll[, c("hit", "false_alarm") := NULL]

# print table
dAll %>%
  kbl(caption = "d-prime for each subject") %>%
  row_spec(0,bold=TRUE) %>%   kable_classic(full_width = F, html_font = "Arial",font_size = 16)
```

## Awareness \* Configuration \* Phase ANOVA on RT saving

```{r 3-way ANOVA on RT saving}
# Construct RT saving on seeing repeated configurations
awardata = RTexp
awardata <- merge(awardata, dAll[, c("Sub", "dprime")], by = "Sub", all.x = TRUE)
awardata_novel <- awardata%>% filter(Configuration == "novel")
setDT(awardata_novel)
setnames(awardata_novel,"RT","RT_novel")

awardata_repeated <- awardata%>% filter(Configuration == "repeated")
setDT(awardata_repeated)
setnames(awardata_repeated,"RT","RT_repeated")

awardata_saving <- awardata_novel %>%
  inner_join(awardata_repeated, by = c("Sub", "Epoch", "dprime")) %>%
  mutate(RT_saving = (RT_novel - RT_repeated)/(0.5*(RT_novel + RT_repeated))) %>%
  dplyr::select(-starts_with("Configuration"), -RT_novel, -RT_repeated)

awardata_saving$Epoch <- as.numeric(awardata_saving$Epoch)

# Fit a mixed-effects model
awarmodel <- lmer(RT_saving ~ Epoch * dprime + (1 | Sub), data = awardata_saving, REML=F)
summary(awarmodel)
# Visualize the interaction
(plot <- ggplot(awardata_saving, aes(x = dprime, y = RT_saving, color = as.factor(Epoch))) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", alpha = 0.1) +
  labs(x = "d'", y = "RT Saving (%)", color = "Epoch") +
  theme_minimal() +
  theme(legend.text = element_text(size = 14), # legend font size
        legend.title=element_text(size=14, face = "bold"), # legend title
        axis.text = element_text(size = 10), # axis text size
        axis.title = element_text(size = 16), # axis title size
  ) )

ggsave("Exp2Slopes.png", plot + theme(plot.background = element_rect(fill = "white")))
```

## Correlation between RT saving in 5th Epoch and d'
```{r second half training correlation}

# Calculate correlation coefficients for each Block
# correlations <- bind_rows(correlations)
awardata_epoch5 <- awardata_saving[Epoch == 5]
correlations <- cor.test(awardata_epoch5$RT_saving, awardata_epoch5$dprime, method = "pearson", use = "pairwise.complete.obs")
p_value <- correlations$p.value
cat("p = ", p_value, "\n")

# print table
correlations <- awardata_epoch5 %>%
  # group_by(Block) %>%
  do(correlation = cor(.$RT_saving, .$dprime, use = "pairwise.complete.obs"))

correlations %>%
  kbl(caption = "r between RT savings and d for last epoch in training phase'") %>%
  row_spec(0,bold=TRUE) %>%   kable_classic(full_width = F, html_font = "Arial",font_size = 16)
```

## Correlation between RT saving in testing phase and d'

```{r testing correlation}
# Calculate correlation coefficients for each Block
# correlations <- bind_rows(correlations)
awardata_epoch5 <- awardata_saving[Epoch == 6]
correlations <- cor.test(awardata_epoch5$RT_saving, awardata_epoch5$dprime, method = "pearson", use = "pairwise.complete.obs")
p_value <- correlations$p.value
cat("p = ", p_value, "\n")

# print table
correlations <- awardata_epoch5 %>%
  # group_by(Block) %>%
  do(correlation = cor(.$RT_saving, .$dprime, use = "pairwise.complete.obs"))

correlations %>%
  kbl(caption = "r between RT savings and d for testing phase'") %>%
  row_spec(0,bold=TRUE) %>%   kable_classic(full_width = F, html_font = "Arial",font_size = 16)
```

## Awareness(binned) \* Configuration \* Phase ANOVA

Categorize subjects whose d-prime \> 0 as aware-group, and the other as
unaware-group.

```{r binned awareness}
# Convert dprime to yes/no
awardata_binned = RTexp
awardata_binned <- merge(awardata_binned, dAll[, c("Sub", "dprime")], by = "Sub", all.x = TRUE)
awardata_binned$Awareness <- ifelse(awardata_binned$dprime > 0, 1, 0)

bxp <- ggboxplot(
  awardata_binned, x = "Awareness", y = "RT",
  color = "Epoch", palette = "jco",
  facet.by = "Configuration", short.panel.labs = FALSE
  )
bxp

res.aov <- anova_test(
  data = awardata_binned, dv = RT, wid = Sub,
  between = Awareness, within = c(Configuration, Epoch)
  )
get_anova_table(res.aov)
```


## Awareness(binned) \* Phase ANOVA on RT saving

```{r binned awareness on RT_saving}
# Convert dprime to yes/no
awardata_binned = RTexp
awardata_binned <- merge(awardata_binned, dAll[, c("Sub", "dprime")], by = "Sub", all.x = TRUE)
awardata_binned$Awareness <- ifelse(awardata_binned$dprime > 0, 1, 0)

awardata_novel <- awardata_binned%>% filter(Configuration == "novel")
setDT(awardata_novel)
setnames(awardata_novel,"RT","RT_novel")

awardata_repeated <- awardata_binned%>% filter(Configuration == "repeated")
setDT(awardata_repeated)
setnames(awardata_repeated,"RT","RT_repeated")

awardata_saving_binned <- awardata_novel %>%
  inner_join(awardata_repeated, by = c("Sub", "Epoch", "Awareness")) %>%
  mutate(RT_saving = RT_novel - RT_repeated) %>%
  dplyr::select(-starts_with("Configuration"), -RT_novel, -RT_repeated)



bxp <- ggboxplot(
  awardata_binned, x = "Awareness", y = "RT",
  color = "Epoch", palette = "jco",
  short.panel.labs = FALSE
  )
bxp

res.aov <- anova_test(
  data = awardata_saving_binned, dv = RT_saving, wid = Sub,
  between = Awareness, within = c(Epoch)
  )
get_anova_table(res.aov)
```


# Z-score Search Time

## Averaged Z-scored RT plots

```{r z-scored}
# Construct new data(tempdata changed)
tempdata[, RT:=scale(RT), by=Sub]
RTnorm <- tempdata[filter==1,
                     .(RT = mean(RT)),
                     keyby = .(Sub,Phase,Configuration,Block, Accuracy)]
RTnorm$Sub <- as.factor(RTnorm$Sub)
RTnorm$Phase <- as.factor(RTnorm$Phase)
RTnorm$Configuration <- as.factor(RTnorm$Configuration)
RTnorm$Block <- as.factor(RTnorm$Block)
RTnorm <- RTnorm %>% filter(Phase != "rt")

##filter4: Binned into epochs
RTnorm <- as.data.table(RTnorm)
RTnorm[, Block := as.numeric(Block)]
RTnorm$Block <- ifelse(RTnorm$Phase == "testing", RTnorm$Block + 20, RTnorm$Block)
RTnorm$Block <- RTnorm$Block - 1
RTnorm[, Epoch := cut(Block, breaks = c(0, 4, 8, 12, 16, 20, 25),
                      labels = c(1, 2, 3, 4, 5, 6), include.lowest = FALSE)]
RTnorm[, Epoch := as.numeric(Epoch)]
RTnorm$Epoch <- as.factor(RTnorm$Epoch)


# within-subject SE
dfwc <- summarySEwithin(RTnorm, measurevar="RT", withinvars=c("Phase","Configuration","Epoch"),
                        idvar="Sub", na.rm=FALSE, conf.interval=.95)
setDT(dfwc)

RT_norm_exp <- dfwc %>% 
  ggplot(aes(x=Epoch, y=RT, 
             colour=factor(Configuration,levels=c("novel","repeated")),
             group=factor(Configuration,levels=c("novel","repeated")))) +
  geom_line(size=1) +
  geom_point(aes(shape=factor(Configuration,levels=c("novel","repeated")),
                 color=factor(Configuration,levels=c("novel","repeated"))),size=2.5) + 
  scale_y_continuous(breaks = seq(-1, 1, by=0.25), limits=c(-1, 1)) +
  geom_errorbar(width=.3, aes(ymin=(RT-se), ymax=(RT+se))) +
  xlab("Epoch") + 
  ylab("Z-Scored RT") +
  scale_color_grey(start = 0.2, end = 0.6) +
  theme_publish() +
  geom_vline(xintercept = 5.4, linetype="dotted") +
  annotate("text", x=3, y=0.75, label= "training", size=15/.pt) + 
  annotate("text", x=6, y=0.75, label= "testing", size=15/.pt) +
  labs(color = "Configuration", shape = "Configuration")+
  labs(title = "Exp 2") +
  theme(legend.position="top",
        aspect.ratio = 0.8)+
  theme(legend.text = element_text(size = 14), # legend font size
        legend.title=element_text(size=14), # legend title
        axis.text = element_text(size = 14), # axis text size
        axis.title = element_text(size = 16), # axis title size
        plot.title = element_text(size = 18,hjust = 0.5), # plot title size
  ) 


plot <- ggarrange(RT_norm_exp, nrow=1,common.legend = TRUE,legend="bottom")
annotate_figure(plot, top = text_grob("",
                                      color = "black", face = "bold", size = 16))

ggsave("Exp2RTStandard.png", plot + theme_minimal() + theme(plot.background = element_rect(fill = "white")))
```


## Z-scored RT ANOVA

```{r RT ANOVA by epoch Z-RT}
# RT by Config across blocks(correct only)
tempdata <- as.data.table(tempdata)
write.csv(tempdata, file = "RTnorm_exp2.csv")
RTexp <- tempdata[filter==1,
                     .(RT = mean(RT)),
                     keyby = .(Sub,Phase,Configuration,Block)]
RTexp$Sub <- as.factor(RTexp$Sub)
RTexp$Phase <- as.factor(RTexp$Phase)
RTexp$Configuration <- as.factor(RTexp$Configuration)
RTexp$Block <- ifelse(RTexp$Phase == "testing", RTexp$Block + 20, RTexp$Block)
RTexp <- RTexp %>% filter(Phase != "rt")
RTexp$Block <- as.factor(RTexp$Block)

##filter4: Binned into epochs
RTexp <- as.data.table(RTexp)
RTexp[, Block := as.numeric(Block)]
RTexp[, Epoch := cut(Block, breaks = c(0, 4, 8, 12, 16, 20, 25),
                        labels = c(1, 2, 3, 4, 5, 6), include.lowest = FALSE)]
RTexp[, Epoch := as.numeric(Epoch)]
RTexp$Epoch <- as.factor(RTexp$Epoch)
RTexp <- RTexp[, .(RT = mean(RT)),
                     keyby = .(Sub,Configuration,Epoch)]
write.csv(RTexp, file = "Exp2AZscoreData.csv")


# repeated ANOVA for RT for main experiment
options(contrasts=c('contr.sum','contr.poly'))
res.aov <- anova_test(
  data = RTexp,
  dv = RT,
  wid = Sub,
  within = c(Configuration,Epoch),
  type = 3,
  effect.size = "pes"
)

# simple effect test
model <- aov(RT ~ Configuration * Epoch + Error(Sub), data = RTexp)
emm <- emmeans(model, ~ Configuration | Epoch)
comparisons <- pairs(emm)
print(comparisons)

# print ANOVA table for main experiment
get_anova_table(res.aov, 
                correction="none") %>%
  kbl(caption = "RT: ANOVA on target's configuration * epoch") %>%
  row_spec(0,bold=TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial",font_size = 16)



```
### Z-scored ANOVA for Training Phase Only
```{r RT ANOVA by subset epoch}
# Filter data for epochs 1 to 5
RTexp_subset <- RTexp[Epoch %in% c(1, 2, 3, 4, 5)]

# Repeated ANOVA for RT for epochs 1 to 5
res.aov_subset <- anova_test(
  data = RTexp_subset,
  dv = RT,
  wid = Sub,
  within = c(Configuration, Epoch),
  type = 3,
  effect.size = "pes"
)


# Bayes Factor Analysis
bf <- anovaBF(RT ~ Configuration, data = RTexp_subset)
bf


# Simple effect test for epochs 1 to 5
model_subset <- aov(RT ~ Configuration * Epoch + Error(Sub), data = RTexp_subset)

emm_subset1 <- emmeans(model_subset, ~ Configuration | Epoch)
comparisons_subset1 <- pairs(emm_subset1)
print(comparisons_subset1)

emm_subset2 <- emmeans(model_subset, ~ Epoch | Configuration)
comparisons_subset2 <- pairs(emm_subset2)
print(comparisons_subset2)

# Print ANOVA table for epochs 1 to 5
get_anova_table(res.aov_subset, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration * epoch (Epochs 1-5)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)
```

### Z-scored ANOVA for 5th Epoch
```{r ANOVA epoch 5 Z-scored}
# Filter data for the 6th epoch
RTexp_epoch5 <- RTexp[Epoch == 5]

# ANOVA for RT for the 6th epoch (main effect of Configuration)
model_epoch5 <- anova_test(
  data = RTexp_epoch5,
  dv = RT,
  wid = Sub,
  within = Configuration,
  type = 3,
  effect.size = "pes"
)


# Cohen's d
repeated_5 <- RTexp_epoch5$RT[RTexp_epoch5$Configuration == "repeated"]
novel_5 <- RTexp_epoch5$RT[RTexp_epoch5$Configuration == "novel"]
cat("RT saving (units):", mean(novel_5) - mean(repeated_5), "\n")


# Print ANOVA table for the 5th epoch with PES
get_anova_table(model_epoch5, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration (Epoch 5)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)


# t test
RTepoch5_wide <- RTexp_epoch5 %>%
  pivot_wider(names_from = Configuration, values_from = RT)
t_test_result <- t.test(RTepoch5_wide$repeated, RTepoch5_wide$novel, paired = TRUE)
formatted_t_test_result <- lapply(t_test_result, function(x) if(is.numeric(x)) sprintf("%.3f", x) else x)

# Print the formatted result
print(t_test_result$statistic)

cohen_d <- t_test_result$statistic/sqrt(length(unique(RTexp_epoch5$Sub[!is.na(RTexp_epoch5$Sub)])))
cat("Cohen's d:", cohen_d, "\n")
```


## Z-scored ANOVA for Transfer Phase
```{r ANOVA transfer phase}
# Filter data for the 6th epoch
RTexp_epoch6 <- RTexp[Epoch == 6]

# ANOVA for RT for the 6th epoch (main effect of Configuration)
model_epoch6 <- anova_test(
  data = RTexp_epoch6,
  dv = RT,
  wid = Sub,
  within = Configuration,
  type = 3,
  effect.size = "pes"
)


# Cohen's d
repeated_6 <- RTexp_epoch6$RT[RTexp_epoch6$Configuration == "repeated"]
novel_6 <- RTexp_epoch6$RT[RTexp_epoch6$Configuration == "novel"]
cat("RT saving (units):", mean(novel_6) - mean(repeated_6), "\n")


# Print ANOVA table for the 6th epoch with PES
get_anova_table(model_epoch6, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration (Epoch 6)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)

# t test
RTepoch6_wide <- RTexp_epoch6 %>%
  pivot_wider(names_from = Configuration, values_from = RT)
t_test_result <- t.test(RTepoch6_wide$repeated, RTepoch6_wide$novel, paired = TRUE)
formatted_t_test_result <- lapply(t_test_result, function(x) if(is.numeric(x)) sprintf("%.3f", x) else x)

# Print the formatted result
print(t_test_result$statistic)

cohen_d <- t_test_result$statistic/sqrt(length(unique(RTexp_epoch6$Sub[!is.na(RTexp_epoch6$Sub)])))
cat("Cohen's d:", cohen_d, "\n")
```


## Z-scored ANOVA for Epoch 5-6
```{r ANOVA two epochs phase}
# Filter data for the 6th epoch
RTexp_compa <- RTexp[Epoch %in% c(5,6)]

# ANOVA for RT for two epochs
model_compa <- anova_test(
  data = RTexp_compa,
  dv = RT,
  wid = Sub,
  within = c(Configuration, Epoch),
  type = 3,
  effect.size = "pes"
)


# Print ANOVA table for epochs 1 to 5
get_anova_table(model_compa, 
                correction = "none") %>%
  kbl(caption = "RT: ANOVA on target's configuration * epoch (Epochs 5-6)") %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 16)
```



# Individual difference
For repeated configuration, we check individual difference in awareness and its impact on RT in testing phase. Specifically, We examined the RT of identified and unidentified repeated configurations separately.

```{r by sub Exp2}
##########################################
# read csv sheet with name
RTlong_exp2 <- read.csv("RT_exp2.csv")
RTnorm_exp2 <- read.csv("RTnorm_exp2.csv")


RTlong_exp2 <- RTlong_exp2 %>%
  mutate(Condition = "tunnel")
RTnorm_exp2 <- RTnorm_exp2 %>%
  mutate(Condition = "tunnel")

# merge data
RTlong <- RTlong_exp2
RTnorm <- RTnorm_exp2
##########################################

## Exp2
filtered_data <- RTlong_exp2 %>%
  filter(Phase == "rt" & Configuration == "repeated")

# Store the aware and unaware config for each Sub
filtered_data <- as.data.table(filtered_data)
filtered_data$configurationNumber = as.numeric(filtered_data$configurationNumber)

aware_data <- filtered_data %>%
  filter(Accuracy == 1) %>%
  dplyr::select(Sub, configurationNumber)

unaware_data <- filtered_data %>%
  filter(Accuracy == 0) %>%
  dplyr::select(Sub, configurationNumber)
  
# Group the data in testing phase
RT_testing <- RTlong %>%
  filter(Phase == "testing")

RT_aware <- RT_testing %>%
  semi_join(aware_data, by = c("Sub", "configurationNumber"))

RT_unaware <- RT_testing %>%
  semi_join(unaware_data, by = c("Sub", "configurationNumber"))

# Merge data based on recognition task
RT_aware <- RT_aware %>%
  mutate(recog = 1)
RT_unaware <- RT_unaware %>%
  mutate(recog = 0)
RT_indi <- rbind(RT_aware, RT_unaware)

# Plot RT saving from awareness
Sub_diff <- RT_indi %>%
  filter(filter == 1) 
RT_aware <- as.data.table(RT_aware)
RT_unaware <- as.data.table(RT_unaware)

RT_aware <- RT_aware[filter == 1, .(RTaware=mean(RT)),keyby=.(Sub)]
RT_unaware <- RT_unaware[filter == 1, .(RTunaware=mean(RT)),keyby=.(Sub)]
RTall <- merge(RT_aware, RT_unaware, by = "Sub", all = TRUE)
RTall <- RTall[!is.na(RTaware) & !is.na(RTunaware)]
RTall[, RTcc := RTunaware - RTaware]

# Creating a scatter plot
(plot <- ggplot(RTall, aes(x = Sub, y = RTcc, color = RTcc)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "blue", high = "red", name = "RT saving") +
  labs(x = "Individuals", y = "RT Saving from Recognition (s)", title = "Individual Differences in Transfer Phase") +
  theme(
    legend.text = element_text(size = 14),  # legend font size and bold
    legend.title = element_text(size = 14, face = "bold"),  # legend title and bold
    axis.text = element_text(size = 10),  # axis text size
    axis.title = element_text(size = 16),  # axis title size
    axis.text.x = element_blank(),  # hide x-axis labels
    plot.title = element_text(size = 18,hjust = 0.5, face = "bold"),
  ))

       
ggsave("Exp2Indi.png", plot + theme(plot.background = element_rect(fill = "white")))       

# ANOVA on RT saving
RT_indi <- as.data.table(RT_indi)
setnames(RT_indi,"recog","Recognition")
RT_indi$Recognition <- as.factor(RT_indi$Recognition)
RT_indi$Sub <- as.factor(RT_indi$Sub)
RT_indi <- RT_indi[, .(RT = mean(RT)),
                     keyby = .(Sub,Recognition)]

options(contrasts=c('contr.sum','contr.poly'))
res.aov <- anova_test(
data = RT_indi,
dv = RT,
wid = Sub,
within = Recognition,
type = 3,
effect.size = "pes"
)

# print ANOVA table for testing
get_anova_table(res.aov, 
                correction="none") %>%
  kbl(caption = "RT: ANOVA on recognition of repeated configuration") %>%
  row_spec(0,bold=TRUE) %>%
  kable_classic(full_width = F, html_font = "Arial",font_size = 16)


recognized_stats <- RT_indi[Recognition == 1, .(mean = mean(RT*1000), se = sd(RT*1000)/sqrt(.N))]
unrecognized_stats <- RT_indi[Recognition == 0, .(mean = mean(RT*1000), sd = sd(RT*1000)/sqrt(.N))]

# Print the results
cat("Stats for recognized old:\n")
print(recognized_stats)

cat("\nStats for unrecognized old:\n")
print(unrecognized_stats)
```
